---
title: "Python Codes"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Python Codes}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  python.reticulate = FALSE,
  collapse = TRUE,
  comment = "#>"
)
```

## Introduction
After generating the data in R and exporting the necessary information, the 
following codes can be used to get the data in python.  

## Generate data in R
If you want to just transform your data in R and export it to python, the `{pins}`
or [reticulate](https://rstudio.github.io/reticulate/index.html) package 
works just fine. But here is how you can do it with `{rpwf}`, which wraps around
`{pins}`.   

```{r}
library(rpwf)
library(pins)
library(recipes)
tmp_dir <- tempdir() # Temp folder
```

* Create a `pins::board_<your board type>` and pass it to `rpwf_connect_db()`.  
* Create a database called `"db.SQLite"` with `rpwf_connect_db()`.  

```{r}
board <- board_temp()
db_con <- rpwf_connect_db(paste(tmp_dir, "db.SQLite", sep = "/"), board)
```

* Create two recipes and pass them to `rpwf_data_set()`  
```{r}
r <- recipe(mpg ~ ., data = mtcars) |>
  step_normalize(all_numeric_predictors()) |>
  rpwf_tag_recipe("r")

r1 <- r |>
  step_YeoJohnson(all_numeric_predictors()) |>
  rpwf_tag_recipe("r1")

d <- rpwf_data_set(r, r1, db_con = db_con)
```

* Write the transformed data, export the metadata to the database, and write the 
board YAML file.  
```{r}
rpwf_write_df(d)
rpwf_export_db(d, db_con)
rpwf_write_board_yaml(board, paste(tmp_dir, "board.yml", sep = "/"))
```

## Get the data  in python
* Import the modules  
* Create a board from the written yml file and a database object  
```{python, eval = FALSE}
from rpwf import database, rpwf
from pathlib import Path

db_path = # <replace with tmp_dir>
board_yml = # <replace with paste(tmp_dir, "board.yml", sep = "/")>

db_obj = database.Base(db_path)
board_obj = database.Board(board_yml)
```

* See all the exported wflow as follows  
```{python, eval = FALSE}
db_obj.all_wflow()

# wflow_id    model_tag   recipe_tag    result_pin_name   model_pin_name
# 1           None        r             None              None
# 2           None        r1            None              None
```

* Pick a `wflow_id`, and create a `rpwf.Wflow` object associated with that 
`wflow_id`
```{python, eval = FALSE}
wflow_id = 2
wflow_obj = rpwf.Wflow(db_obj, board_obj, wflow_id)
```

* Finally create a `rpwf.TrainDf` object and use the `get_df_X` and `get_df_y` 
methods to get the train `pandas.DataFrame` and response `pandas.Series`  
```{python, eval = FALSE}
df_obj = rpwf.TrainDf(db_obj, board_obj, wflow_obj)
X, y = df_obj.get_df_X(), df_obj.get_df_y()
```

